---
title: "BCFR bc confirmation"
author: "Ziyi Zhao"
date: "2/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

Since we only use data from registries in Canada and US, we would not use any tables or data about Austrilia.

# Breast Confirmation
* all variables are included
```{r}
# import the table of breast confirmation
## bc_confirm <- read_csv("...") %>% janitor::clean_names() %>% as_tibble()

# filter out registries in US and Canada
## bc_confirm <- filter(bc_confirm, center_no %in% c(1,2,3,5,6))

# also need to filter out missing and not asked
## bc_confirm_miss <- filter(bc_confirm, 
##                           center_no==9|person_id==999999999999|tumor_no==99|
##                           source_br==9|br_dx_conf==9|br_tiss==9)
## bc_confirm_nask <- filter(bc_confirm,
##                          center_no==8|person_id==888888888888|tumor_no==88|
##                           source_br==8|br_dx_conf==8|br_tiss==8)

```

Problem: 
* what does 'test site' mean in varaible center_no?

* What shall we do with variable 'Source_BR'? Does source of breast cancer matter?

* What can we do with missing and not asked subjects?

* For variable breast surgeries, should we only keep subject without breast surgeries? 




