---
title: "BCFR breast epi"
author: "Ziyi Zhao"
date: "2/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)


```

# Breast Epi

* primary key is personal id (unique), so every tables we created need to carry personal id.

```{r}
# import the breast_epi
## bc_epi <- read_csv("...") %>% janitor::clean_names()

# filter out subjects in US and Canndian registries
## bc_epi <- filter(bc_epi, center_no %in% c(1,2,3,5,6))

# select variables about age
## bc_epi_age <- select(bc_epi,person_id,dt_interview,age_interview)

# select varaibles about education
## bc_epi_edu <- select(bc_epi,person_id,education)

# select variables about marital status
## bc_epi_marstat <- select(bc_epi,person_id,marstat)

# select variables about race
## bc_epi_psid <- select(bc_epi,person_id)
## bc_epi_race <- select(bc_epi,contains("race")) %>% cbind(bc_epi_psid)

# select variables about height and BMI
## bc_epi_mbi <- select(bc_epi,person_id,height,weight) %>% 
##                  mutate(ht_m = height*2.54/100,
##                         wt_kg = weight*0.453,
##                         bmi = wt_kg/(ht_m^2))

# select variables about alcohol
## bc_epi_bwl <- select(bc_epi,person_id,beer_wk,wine_wk,liquor_wk)
## bc_epi_alcoh <- select(bc_epi,contains("alcoh")) %>% 
##                    cbind(bc_epi_bwl)

# select variable about cigarette smoking
## bc_epi_cig <- select(bc_epi,starts_with("cig")) %>% cbind(bc_epi_psid)

# select varaibles about physical activities
## bc_epi_strn <- select(bc_epi,starts_with("stren"))
## bc_epi_mod <- select(bc_epi,starts_with("mod"))
## bc_epi_phyc <- cbind(bc_epi_strn,bc_epi_mod,bc_epi_psnid)

# 




```

Problem about data:

* why are there four variables about race? (their questions are exactly same, so why repeated?)

* Why is there a variable about race in text? new variable for race.

* The physical activity is divided into two parts: strenuous and moderate.How do we deal with these two? 

* The units of physical activity also separate into hours/week and month/year. Which unit should we select? 
